---
title: "DATA607FinalProject"
author: "Daniel Hong"
date: "May 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE, cache=FALSE}
# DO NOT REMOVE
# THIS IS FOR SETTING SOME PLOTTING PARAMETERS SO THAT YOUR PLOTS DON'T TAKE UP TOO MUCH SPACE
# IF YOU WOULD LIKE TO CHANGE THESE, SEE HELP FILES ON THE par() FUNCTION
# OR ASK FOR HELP
library(knitr)
## set global chunk options
opts_chunk$set(fig.path='figure/manual-', cache.path='cache/manual-', fig.align='center', fig.show='hold', par=TRUE)
## tune details of base graphics (http://yihui.name/knitr/hooks)
knit_hooks$set(par=function(before, options, envir){
if (before && options$fig.show!='none') par(mar=c(4,4,.2,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
})
```

  Load required libraries
```{r, eval=TRUE}
library(devtools)
library(finstr)
library(XBRL)
library(htmlTable)
library(ggplot2)
library(tidyr)
library(dplyr)
library(curl)
```

### Introduction

  Publicly traded coportation are required to file quarterly (10-Q) and annual statements (10-K) with the Securities    and Exchange Commission (SEC). While there are many other filings filed with the SEC we will focus on these two       versions in this project.
  
  Individul and professional investors look through these reports for nuggets of information to help make decisions.    Over the years parts of the reports have become automatable through technological advances and formatting. We will    look at how the formatting has helped us analyze certain parts of the filing. The purpose is to help save time        looking at the financials so we can spend more time analyzing the text. Improvements have been made in automated      text analysis but it is still generally accepted that a human's contribution is still very important.
    
  What is XBRL? XBRL is the open international standard for digital business reporting, managed by a global not for     profit consortium, XBRL International. We are  committed to improving reporting in the public interest. XBRL is used   around the world, in more than 50 countries. Millions of XBRL documents are created every year, replacing older,      paper-based reports with more useful, more effective and more accurate digital versions.
  
  We will look at possibly the most popular company for the past decade Apple (AAPL), a wall sreet darling. Each        company is assigned a CIK number, in this case 0000320193. Another important identifier is the Standard Industry      Classification (SIC) code, SIC:3571 Electronic Computers. This system aims to classify industries by a government     accepted 4 digit code, these codes are utilized in many ways. 
  
### Data
  Parse SEC filings 10-Ks and retrieve the income statement
```{r}
library(pander)

xbrl_url2014 <- "http://edgar.sec.gov/Archives/edgar/data/320193/000119312514383437/aapl-20140927.xml"
xbrl_url2013 <- "http://edgar.sec.gov/Archives/edgar/data/320193/000119312513416534/aapl-20130928.xml"
old_o <- options(stringsAsFactors = FALSE)
xbrl_data_aapl2014 <- xbrlDoAll(xbrl_url2014)
xbrl_data_aapl2013 <- xbrlDoAll(xbrl_url2013)
options(old_o)

st2013 <- xbrl_get_statements(xbrl_data_aapl2013)
st2014 <- xbrl_get_statements(xbrl_data_aapl2014)

income2013 <- st2013$StatementOfIncome
income2014 <- st2014$StatementOfIncome
is_all <- merge(income2013, income2014)
print(is_all, html = TRUE, big.mark = ",", dateFormat = "%Y")
```

  Retrieve the balance sheet
```{r}
balance_sheet2013 <- st2013$StatementOfFinancialPositionClassified
balance_sheet2014 <- st2014$StatementOfFinancialPositionClassified
bs_all <- merge(balance_sheet2013, balance_sheet2014)
print(bs_all, html = TRUE, big.mark = ",", dateFormat = "%Y")
```

  Retrieve the statement of cash flows
```{r}
cashFlow2013 <- st2013$StatementOfCashFlowsIndirect
cashFlow2014 <- st2014$StatementOfCashFlowsIndirect
cf_all <- merge(cashFlow2013, cashFlow2014)
print(cf_all, html = TRUE, big.mark = ",", dateFormat = "%Y")
```

  Merge all statements and check if the balance sheets are merged
```{r}
st_all <- merge(st2013, st2014)
balance_sheet <- st_all$StatementOfFinancialPositionClassified
bs_simple <- expose(balance_sheet,
                    `Current Assets` = "AssetsCurrent",
                    `Noncurrent Assets` = other("Assets"),
                    `Current Liabilities` = "LiabilitiesCurrent",
                    `Noncurrent Liabilities` = other(c("Liabilities", "CommitmentsAndContingencies")),
                    `Stockholders Equity` = "StockholdersEquity")
print(bs_simple, html = TRUE, big.mark = ",", dateFormat = "%Y")
```

### Plots
```{r}
plot_double_stacked_bar(bs_simple)
```

```{r}
plot_double_stacked_bar(bs_simple, by_date = FALSE)
```

```{r}
bs_simple_prop <- proportional(bs_simple)
plot_double_stacked_bar(bs_simple_prop)
```

### Data Analysis
  The current ratio is a liquidity and efficiency ratio that measures a firm's ability to pay off its short-term        liabilities with its current assets. The current ratio is an important measure of liquidity because short-term        liabilities are due within the next year.
```{r}
balance_sheet %>% transmute(
  date = endDate, 
  CurrentRatio = AssetsCurrent / LiabilitiesCurrent
)
```

  Days sales outstanding (DSO) is a calculation used by a company to estimate their average collection period. It is a   financial ratio that illustrates how well a company's accounts receivables are being managed.
```{r}
merge(balance_sheet, st_all$StatementOfIncome ) %>% calculate( digits = 2,

    .AccountReceivableLast = lag(AccountsReceivableNetCurrent),
    .AccountReceivableAvg = (.AccountReceivableLast + AccountsReceivableNetCurrent)/2,

    DaysSalesOutstanding = .AccountReceivableAvg / SalesRevenueNet * 365 
  )
```

  Define Calculation - Profit Margins
```{r}
profit_margins <- calculation(

  Gross_Margin = 
    (SalesRevenueNet -  CostOfGoodsAndServicesSold) / SalesRevenueNet,

  Operating_Margin =
    OperatingIncomeLoss / SalesRevenueNet,

  Net_Margin = 
    NetIncomeLoss / SalesRevenueNet

)

income2013 %>% calculate(calculations = profit_margins, digits = 2)
income2014 %>% calculate(calculations = profit_margins, digits = 2)
```

  Define Calculation - Accounts Receivable / SalesRevenue
```{r}
merge(balance_sheet, st_all$StatementOfIncome ) %>% calculate( digits = 2,

    .AccountReceivableLast = lag(AccountsReceivableNetCurrent),
    .AccountReceivableAvg = (.AccountReceivableLast + AccountsReceivableNetCurrent)/2,

    quality = .AccountReceivableLast / SalesRevenueNet
  )
```

  Over the 3 periods, Apple's ratio of Accounts Receivable to sales remained consistent year over year. We could not    confirm their gift card liability to sales ratio but suspect that it remained constant as well. These two metrics     indicates that they most probably did not overstate revenues. Aaple has been fortunate to increase revenues, albeit   over the last 3 years, less fruitful companies could manipulate earnings by improperly recognizing deferred revenue.
  When we take a closer look at Apple's margins, it is unlikely that they overproduced to spread costs over a larger    base of goods to lower COGS. The profit margin declines are most likely due to new product designs. It is not         included here, but Apple did in fact increase margins and the stock price reflected this margin expansion.
  
  Another potential red flag is a significant change in revenues attributed to the product mix. In the table below we   see consistency across Apple's product and service segments. The fluctuations in the chart are due to seasonality,    not actual material changes in the product mix.
  
  Apple's SG&A expense increased each year and it appears they the course and continued to investment for future        earnings. The accounting quality of their expense reporting indicates that future earning should be consistent with   current earnings.
  
  Load industry data and run correlations
```{r, eval=TRUE}
capital_structure <- read.csv(curl("https://raw.githubusercontent.com/danielhong98/DATA607/af5de57f3e226c48cb9003897b9673c93752e24a/606FinalProjectData.csv"), header = TRUE)
kable(head(capital_structure,10))

# correlation of Stock price standard deviation to revenue growth
cor(capital_structure$StdevStock,capital_structure$RevenueGrowth)

# correlation of cost of Equity to revenue growth
cor(capital_structure$CostofEquity,capital_structure$RevenueGrowth)
```

  Use the lm function to fit the linear model
```{r, eval=TRUE}
# linear model using DebtPct
m1 <- lm(RevenueGrowth ~ StdevStock, data = capital_structure)

# linear model using CostofCapital
m2 <- lm(RevenueGrowth ~ CostofEquity, data = capital_structure)
```

  The first argument in the function lm is a formula that takes the form y ~ x. We want the linear model of percentage   standard deviation for m1 and Costofequity for m2 as a function of revenue growth.
  
  The output of lm is an object that contains the information we need from the linear model
  that was just fit. We summarize the information:
```{r, eval=TRUE}
summary(m1)

summary(m2)
```

  In the "Coefficients" table above; the first column displays the linear model's y-intercept and the
  coefficient of Stdev for m1 and CostofEquity for m2. With this table, we can write down the least squares             regression line for the linear model:
  
  for m1:  
  \[
  \hat{y} = 0.06552 + 0.04845 * StdevStock
  \]

  for m2:  
  \[
  \hat{y} = 0.10892 - 0.20127 * CostofCapital 
  \]
  
  The $R^2$ value represents the proportion of variability in the response variable that is explained by the            explanatory variable. For m1, 0.84% of the variability in RevenueGrowth is explained by StdevStock. For m2, 0.20%     variablility is explained by CostofCapital.
   
  We see a positive relationship in StdevStock, for every 1 percent of DebtPct we would see an increase of 0.04845      increase in revenue growth. We see a negative relationship in CostofCapital, for every 1 percent of CostofCapital we   would see a decrease of 0.20127 in revenue growth.
  
  To assess whether the linear model is reliable, we need to check for (1) linearity, (2) nearly normal residuals, and   (3) constant variability.
```{r, eval=TRUE}
plot(m1$residuals ~ capital_structure$StdevStock)
abline(h = 0, lty = 3)

plot(m2$residuals ~ capital_structure$CostofEquity)
abline(h = 0, lty = 3)
```
  
  The function abline plots a line based on its slope and intercept.
```{r}
plot(capital_structure$RevenueGrowth ~ capital_structure$StdevStock)
abline(m1)

plot(capital_structure$RevenueGrowth ~ capital_structure$CostofEquity)
abline(m2)
```

### Conclusion
  We looked at a process to help with our analytic approach, but this just scratches the surface. We can look at many   other indicators to examine the quality of Property, Plant and Equipment (PPE), Quality of credit provisions and tax   reporting, additional profitability measures (return on assets, return on capital, return on equity, return on net    operating assets), and asset turnover. The statistical analysis is included for illustration. It gives an overall     assessment for a basket of companies, which Apple is included but futher analysis can be done here as well.
  
### References
  https://www.xbrl.org
  https://www.osha.gov/pls/imis/sicsearch.html
  https://cran.r-project.org/web/packages/XBRL
  https://github.com/bergant/XBRLFiles
  http://people.stern.nyu.edu/adamodar/New_Home_Page/data.html